---
title: "Inventory QA/QC - Notebook"
author: "Vincyane Badouard"
date: "`r Sys.Date()`"
output: html_notebook
---
```{r display_help function, echo = F, results = 'hide'}
display_help <- function(subject, pkg = NULL) {
  tc <- textConnection("s", "w", local = TRUE)
  tools:::Rd2HTML(utils:::.getHelpFile(help(subject, (pkg))), out = tc)
  knitr::asis_output(htmltools::htmlPreserve(s))
}
```

# Install R packages
```{r install pkgs}
devtools::install_github("Alliance-for-Tropical-Forest-Science/DataHarmonization", build_vignettes = TRUE) # To format inventory data
devtools::install_github("VincyaneBadouard/TreeData", build_vignettes = TRUE) # to detect and/or correct inventory data errors 
install.packages(c("readr"))
```

# Load R packages
```{r setup}
library(TreeData)
library(readr)
library(data.table)
library(ggplot2)
```

# Load your tree inventory data
Detail: Dataset (data.frame or data.table)
```{r, message=F}
Data <- read_csv("")
```

# Load the shapefile of your plot
Detail: plot polygon with a crs (sf)
```{r}
PlotPolygon <- st_read("")
```

# Format inventory
```{r}
DataHarmonization::run_app()
```

# QA/QC
All the detection/correction functions of the package:
- Check missing values:
DetectMissingValues()

- Check of the unique association of the IdTree/IdStem with Plot-Subplot-TreeFieldNum, at the site scale:
DetectDiffTreeIDAssociation()

- Check duplicated IdTree/IdStem in a census:
DetectDuplicatedStemID()

- Check the stem number coherence:
DetectStemNbrIncoherence()

- Check invariant coordinates per IdTree/IdStem:
DetectStemVariantCoord()

- Check for trees outside the plot: 
DetectTreesOutBotanicalCorrection() 

- Check/correct the tree life status evolution over the censuses: 
StatusCorrection()  

- Taper correction for buttressed trees: 
TaperCorrection()  

- Check/correct the tree diameter values:
DiameterCorrection()  

- Check/correct missing recruited trees:
RecruitmentCorrection()  

Visualisation functions:
LifeStatusCorrectionPlot()
DiameterCorrectionPlot()
RecruitmentCorrectionPlot() (TODO)

**Load the example dataset stored in the package**
```{r}
data("TestData")
Data <- TestData

data("PlotPolygon")
```

# Check duplicated rows
```{r}
# Are there duplicates rows?
anyDuplicated(Data) != 0
```

## Delete duplicated rows
```{r}
Data <- unique(Data)
```

# Check missing values
```{r}
display_help("DetectMissingValues")
```
```{r}
Data <- DetectMissingValues(
  Data,
  Vars = c("Plot", "Subplot", "Year", "TreeFieldNum", "IdTree", "IdStem", "Diameter","POM", "HOM", "TreeHeight", "StemHeight", "XTreeUTM", "YTreeUTM", "Family", "Genus", "Species", "VernName"),
  MeasVars = c("Diameter", "HOM", "TreeHeight", "StemHeight")
)
```

## Display rows with missing values
```{r}
# The detected errors
unique(Data[Comment != "", Comment])

Data[grepl("Missing value", Comment), ]

Data[grepl("cannot be 0", Comment), ]
```

# Check the unique association of the IdTree/IdStem with Plot-Subplot-TreeFieldNum, at the site scale
```{r}
display_help("DetectDiffTreeIDAssociation")
```
```{r}
Data <- DetectDiffTreeIDAssociation(Data)
```

# Check duplicated IdTree/IdStem in a census
```{r}
display_help("DetectDuplicatedStemID")
```
```{r}
Data <- DetectDuplicatedStemID(Data)
```

# Check the stem number coherence
```{r}
display_help("DetectStemNbrIncoherence")
```
```{r}
Data <- DetectStemNbrIncoherence(Data)
```

## Display trees without stem no.1
```{r}
Data[grepl("Trees without stem no.1", Comment), ]
```

# Check invariant coordinates per IdTree/IdStem
```{r}
display_help("DetectStemVariantCoord")
```
```{r}
Data <- DetectStemVariantCoord(Data)
```

## Display trees with different coordinates
```{r}
Data[grepl("Different coordinates", Comment), ]
```

# Check for trees outside the subplot
```{r}
display_help("DetectTreesOut")
```
```{r}
Data <- DetectTreesOut(Data, PlotPolygon) # it plots the trees in and out the plot also
```

## Display rows with dectected errors
```{r}
# The detected errors
unique(Data[Comment != "", Comment])

Data[Comment != "", ]
```

# Detect botanical errors
```{r}
display_help("DetectBotanicalErrors")
```
```{r}
Data <- DetectBotanicalErrors(Data)
```

## Display rows with botanical errors
```{r}
Data[grepl("aceae", Comment), ]
Data[grepl("Special characters", Comment), ]
Data[grepl("Different botanical informations", Comment), ]
```

# Corect/update botanical informations
```{r}
print("TODO: HERE USE BIOMASS PACKAGE")
```


# Detect multistem
Detect multistem in trees inventory data from common coordinates and botanical name
```{r}
display_help("DetectMultistem")
```
```{r}
Data <- DetectMultistem(Data)
```
## Display potential multistem trees
```{r}
Data[grepl("Potential multistem", Comment), .(IdTree, IdStem, ScientificName)]
```

# Check/correct the tree life status evolution over the censuses
```{r}
display_help("StatusCorrection")
```
```{r}
Data <- StatusCorrection(
  Data,
  InvariantColumns = c("Site", "Genus_TreeDataCor", "Species_TreeDataCor",
                       "Family_TreeDataCor", "ScientificName_TreeDataCor"),
  DeathConfirmation = 2,
  UseSize = FALSE,
  AddRowsForForgottenCensuses = TRUE,
  DetectOnly = FALSE,
  RemoveRBeforeAlive = FALSE,
  RemoveRAfterDeath = FALSE
)
```
## Display tree life status correction
```{r}
corrID <- unique(Data[LifeStatus_TreeDataCor != LifeStatus,]$IdStem)

Data[IdStem %in% corrID, 
     .(IdStem, Year, LifeStatus, LifeStatus_TreeDataCor, Comment)]
```

## Plot tree life status corrections
```{r}
display_help("LifeStatusCorrectionPlot")
```

```{r}
# To display the plots in the Rstudio session:
LifeStatusCorrectionPlot(Data,
                         OnlyCorrected = TRUE,
                         Rmarkdown = TRUE,
                         SeveralWindows = TRUE)

# To download as pdf:
# pdf("LifeStatusCorrectionPlot_TestData.pdf", width = 25, height = 10)
# LifeStatusCorrectionPlot(Data, OnlyCorrected = TRUE, SeveralWindows = FALSE)
# dev.off()
```

# Taper correction for buttressed trees
Transform the tree diameter measured at a given height into the diameter corresponding to the default measurement height (HOM), using an allometry
```{r}
display_help("TaperCorrection")
```

```{r}
Data <- TaperCorrection(
  Data,
  DefaultHOM = 1.3,
  TaperParameter = function(DAB, HOM) 0.156 - 0.023 * log(DAB) - 0.021 * log(HOM),
  TaperFormula = function(DAB, HOM, TaperParameter, DefaultHOM) DAB/(exp(-TaperParameter * (HOM - DefaultHOM))),
  DetectOnly = FALSE
)
```

# Diameter correction
```{r}
display_help("DiameterCorrection")
```
```{r}
Data <- DiameterCorrection(
  Data,
  KeepMeas = c("MaxHOM", "MaxDate"),
  DefaultHOM = 1.3,
  MaxDBH = 500,
  PositiveGrowthThreshold = 5,
  NegativeGrowthThreshold = -2,
  Pioneers = NULL,
  PioneersGrowthThreshold = 7.5,
  WhatToCorrect = c("POM change", "punctual", "shift"),
  CorrectionType = c("linear", "individual"), # ,"phylogenetic hierarchical"
  DBHRange = 10,
  MinIndividualNbr = 5,
  OtherCrit = NULL,
  Digits = 2L,
  DBHCorForDeadTrees = TRUE,
  coef = 0.9,
  DetectOnly = FALSE)
```

# Plot diameter corrections
```{r}
display_help("DiameterCorrectionPlot")
```
```{r}
# To display the plots in the Rstudio session:
DiameterCorrectionPlot(Data,
                       OnlyCorrected = TRUE,
                       CorCol = "Diameter_TreeDataCor",
                       Rmarkdown = TRUE,
                       SeveralWindows = TRUE)

# To download as pdf:
# pdf("DiameterCorrectionPlots_TestData.pdf", width = 25, height = 10)
# DiameterCorrectionPlot(Data, OnlyCorrected = TRUE, SeveralWindows = FALSE)
# dev.off()
```


# Check/correct missing recruited trees
```{r}
display_help("RecruitmentCorrection")
```
```{r}
Data <- RecruitmentCorrection(
  Data,
  KeepMeas = c("MaxHOM", "MaxDate"),
  MinDBH = 10,
  MinDBHLimit = 5,
  PositiveGrowthThreshold = 5,
  InvariantColumns = c("Site", "Genus_TreeDataCor", "Species_TreeDataCor",
                       "Family_TreeDataCor", "ScientificName_TreeDataCor"),
  DetectOnly = FALSE)
```

# Plot corrected recruits
```{r}
display_help("RecruitmentCorrectionPlot")
```

```{r}
# To display the plots in the Rstudio session:
RecruitmentCorrectionPlot(Data,
                          OnlyCorrected = TRUE,
                          CorCol = "Diameter_TreeDataCor",
                          Rmarkdown = TRUE,
                          SeveralWindows = TRUE)
# To download as pdf:
# pdf("RecruitmentCorrectionPlot_TestData.pdf", width = 25, height = 10)
# RecruitmentCorrectionPlot(Data, OnlyCorrected = TRUE, SeveralWindows = FALSE)
# dev.off()
```

# Plot initial and corrected diameter distribution of the stand
```{r}
DistributionPlot(Data,
                 InitialCol = "Diameter",
                 CorCol = "Diameter_TreeDataCor")
```

# Plot initial and corrected diameter distribution of the stand
```{r}
DistributionPlot(Data,
                 InitialCol = "TreeHeight")
```

# Plot the height - diameter relationship of the stand
```{r}
HeightVSDiameterPlot(Data, "TreeHeight", "Diameter_TreeDataCor")
```

# Reverse inventory formating
```{r}
Data <- ReversedRequiredFormat(Data)
```

