---
title: "TreeData"
subtitle: "Forest Inventories Harmonization & Correction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{TreeData}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = F, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", echo = T, message = T, warning = F, cache = T
)
```

# Load TreeData and datasets
**Install TreeData**
```{r, eval=F}
devtools::install_github("VincyaneBadouard/TreeData", build_vignettes = TRUE)
```
**Load the packages**
```{r setup}
library(TreeData)
library(knitr)
library(kableExtra)
library(shiny)
library(shinydashboard)
library(shinyjs)
library(shinyWidgets)
library(data.table)
library(ggplot2)
```

# Open the shiny app
```{r, eval = F}
# TreeData::RunApp() # doest' work, why?
shiny::runGitHub( "VincyaneBadouard/TreeData", subdir = "inst/app") # data.tree, stringdist
```


**Load the example dataset stored in the package**
```{r}
data("TestData")
```

# Metadata of the output columns of the TreeData package

- *Comment* (character): error type informations  

**Bota**  
- *Family_TreeDataCor* (character): corrected *Family* name  
- *FamilyCorSource* (character): source of the *Family* correction  
- *Genus_TreeDataCor* (character): corrected *Genus* name  
- *Species_TreeDataCor* (character): corrected *Species* name  
- *BotanicalCorrectionSource* (character): source of the *Genus* and *Species*
correction  
- *ScientificName_TreeDataCor* (character): corrected *ScientificName*  
- *VernName_TreeDataCor* (character): completed if information available at *IdTree*
level. 

**Life Status**  
- *LifeStatus_TreeDataCor* (logical): corrected stem life status.  

**Taper**  
- *TaperDBH_TreeDataCor* (numeric): corrected trees diameter at default *HOM*  

**Diameter**  
- *Diameter_TreeDataCor* (numeric): corrected trees diameter at default *HOM*  
- *DiameterCorrectionMeth* (character): diameter correction methode = "*taper*"/"*local linear regression*"/"*weighted mean*"
/phylogenetic hierarchical("*species*"/"*genus*"/"*family*"/"*stand*")/"*shift realignment*"/"*Same value*".  
- *POM_TreeDataCor* (factor): *POM* value at which the corrected diameters are proposed. Corresponds to the 1st *POM* value at which the stem was measured  
- *HOM_TreeDataCor* (numeric): *HOM* value at which the corrected diameters are proposed. Corresponds to the *DefaultHOM* if *Taper* correction applied. If not, corresponds to the 1st *HOM* value at which the stem was measured  

**Recruitment**  
- *CorrectedRecruit* (logical): TRUE: the row was added to represent the year when the stem was larger than the minimum diameter, but absent from the dataset. FALSE: original dataset row.  

# All the correction functions of the package
GeneralErrorsDetection()  
BotanicalCorrection()  
StatusCorrection()  
TaperCorrection()  
DiameterCorrection()  
RecruitmentCorrection()  

# The corrections illustration functions of the package
StatusCorrectionPlot()  
DiameterCorrectionPlot()  
BotanicalCorrectionPlot()  

# Arguments presentation

# Decomposed corrections

## General errors detection
Detect errors
- Remove **duplicated rows**
- Check **missing value** in
X-YTreeUTM/PlotArea/Plot/Subplot/Year/TreeFieldNum/
IdTree/IdStem/Diameter/POM/HOM/Family/Genus/Species/VernName
- Check **missing value** (NA/0) in the measurement variables: "Diameter",
"HOM", "TreeHeight", "StemHeight"
- Check of the **unique association of the IdTree with plot, subplot**
**and TreeFieldNum** (at the site scale)
- Check **duplicated IdTree/IdStem** in a census (at the site scale)
- Check for trees **outside the subplot** (not implemented yet)
- Check **invariant coordinates per IdTree/IdStem**
- Check **fix Plot and Subplot number** (not implemented yet)

```{r}
Rslt <- GeneralErrorsDetection(TestData)

# The detected errors
unique(Rslt[Comment != "", Comment])

Rslt[Comment != "", ]
```

## Botanical identification correction

- No special characters (typography)
- No family name in the Genus and Species columns (the suffix "aceae" is
specific to the family name.
- Correct spelling of botanical names (*Taxonstand or WorldFlora*)
- Family & Scientific names match (*BIOMASS::getTaxonomy or WorldFlora*)
- Update the scientific botanical names with the current phylogenetic
classification
- Check **invariant botanical informations per IdTree** (1 IdTree = 1 family,
1 scientific and 1 vernacular name)

```{r, eval=F}
Rslt <- BotanicalCorrection(TestData, Source = "TPL", WFOData = NULL)
```

## Life status
- if UseSize argument chosen : if Diameter != NA -> Alive
If (the value in bold is modified by the value given after the arrow):
(the ">" gives the chronological order of the sequence)

- *Dead* > Alive -> NA
- add rows for the forgotten censuses between 2 'Alive' if chosen
- Alive > *Dead*/*NA* > Alive -> Alive
- Alive > *NA* > Dead -> NA
- Alive > *Dead* > NA -> Dead
- Alive > *NA* > *NA*:
if DeathConfirmation > unseens -> NA
if DeathConfirmation =< unseens -> Dead

```{r}
Status_corr <- StatusCorrection(TestData,
                                InvariantColumns = c("Site",
                                                     "Genus",
                                                     "Species",
                                                     "Family",
                                                     "ScientificName"),
                                DeathConfirmation = 2,
                                UseSize = FALSE,
                                AddRowsForForgottenCensuses = TRUE,
                                DetectOnly = FALSE,
                                RemoveRBeforeAlive = TRUE,
                                RemoveRAfterDeath = TRUE)

Status_corr[LifeStatus!=LifeStatus_TreeDataCor]

# pdf("LifeStatusCorrectionPlots.pdf", width = 25, height = 10)
LifeStatusCorrectionPlot(Status_corr, OnlyCorrected = T, SeveralWindows = F)
# dev.off()
```


## Taper correction
Transform the tree diameter measured at a given height into the diameter
corresponding to the default measurement height (HOM), using an allometry.

```{r}
Data <- data.table(IdStem = "A",
                   ScientificName = "Tree",
                   Year = c(1998, 2008, 2016, 2017, 2018, 2019, 2021),
                   IdCensus = factor(c(1998, 2008, 2016, 2017, 2018, 2019, 2021), ordered = TRUE),
                   Diameter = c(19, 19, 21.4, 22.6, 23.1, 23.1, 23.6),
                   HOM = c(1.30, 3.25, 3.25, 3.25, 3.25, 3.25, 3.25))

Rslt <- TaperCorrection(Data, DefaultHOM = 1.3,
                        TaperParameter = function(DAB, HOM) 0.156 - 0.023 * log(DAB) - 0.021 * log(HOM),
                        
                        TaperFormula = function(DAB, HOM, TaperParameter, DefaultHOM) DAB / (exp(- TaperParameter*(HOM - DefaultHOM))))

DiameterCorrectionPlot(Rslt, CorCol = "TaperDBH_TreeDataCor")
```

```{r, eval = F}
Rslt <- TaperCorrection(Status_corr, DefaultHOM = 1.3,
                        TaperParameter = function(DAB, HOM) 0.156 - 0.023 * log(DAB) - 0.021 * log(HOM),
                        
                        TaperFormula = function(DAB, HOM, TaperParameter, DefaultHOM) DAB / (exp(- TaperParameter*(HOM - DefaultHOM))))

DiameterCorrectionPlot(Rslt, CorCol = "TaperDBH_TreeDataCor", OnlyCorrected = TRUE, SeveralWindows = T)

```

## Diameter correction

```{r}
Diam_corr <- DiameterCorrection(
  Status_corr,
  KeepMeas = c("MaxHOM", "MaxDate"),
  
  DefaultHOM = 1.3,
  MaxDBH = 500,
  PositiveGrowthThreshold = 5,
  NegativeGrowthThreshold = -2,
  
  Pioneers = NULL,
  PioneersGrowthThreshold = 7.5,
  
  WhatToCorrect = c("POM change", "punctual", "shift"),
  CorrectionType = "individual",
  
  Digits = 1L,
  
  DBHCorForDeadTrees = FALSE,
  
  coef = 0.9,
  
  DetectOnly = FALSE)

```

```{r, eval = F}
# pdf("DiameterCorrectionPlots_TestData2.pdf", width = 25, height = 10)
DiameterCorrectionPlot(Diam_corr, OnlyCorrected = TRUE, SeveralWindows = F)
# dev.off()
```

## Recruitment

```{r}
Rslt <- RecruitmentCorrection(Diam_corr,
                              KeepMeas = c("MaxHOM", "MaxDate"),
                              MinDBH = 10,
                              PositiveGrowthThreshold = 5,
                              InvariantColumns = c("Site",
                                                   "Genus_TreeDataCor",
                                                   "Species_TreeDataCor",
                                                   "Family_TreeDataCor",
                                                   "ScientificName_TreeDataCor"),
                              DetectOnly = FALSE)

IdCorr <- Rslt[CorrectedRecruit %in%  TRUE, IdStem]
TreesCorr <- Rslt[IdStem %in% IdCorr, .(IdStem,
                                        Year,
                                        Diameter,
                                        Diameter_TreeDataCor,
                                        CorrectedRecruit)]

ggplot(TreesCorr) +
  aes(x = Year, y = Diameter_TreeDataCor) +
  geom_line(linewidth = 0.5, colour = "#112446") +
  geom_point(shape = "circle", size = 1.5, mapping = aes(color = CorrectedRecruit)) +
  theme_minimal() +
  facet_wrap(vars(IdStem), scales = "free")
```


# Interesting other variables
## Mortality rate
## Recruitment rate
## Growth rate
## Basal area
## Wood density
## Biomass
## Tree height
## Disturbance intensity
